---
title: "Assignment-b1"
output: github_document
date: "2022-11-04"
---

```{r, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(datateachr)
library(testthat)
library(dplyr)
library(lubridate)
```

# Exercise 1 & 2

- Exercise 1: In this exercise, you’ll be making a function and fortifying it. The function need not be complicated. The function need not be “serious”, but shouldn’t be nonsense.

- Exercise 2: In the same code chunk where you made your function, document the function using roxygen2 tags. Be sure to include:
1. Title.
2. Function description: In 1-2 brief sentences, describe what the function does.
3. Document each argument with the @param tag, making sure to justify why you named the parameter as you did.
4. (Justification for naming is not often needed, but we want to hear your reasoning.)
5. What the function returns, using the @return tag.

```{r, message=FALSE}
#' @title count_by_time function
#' 
#' @details Given a Date column, count the occurrence of a date unit specified by floor_unit. This could be useful when finding the pattern between values and specified time frames.
#' 
#' @param dataset, the dataset to be analyzed
#' 
#' @param date_column, name of the Date column on which to be applied the operation
#' 
#' @param floor_unit, the date unit to be counted
#' 
#' @return a N X 2 dataframe with column "time" and "count", which represents the list of date unit values and the number of their occurrence, respectively.  
#' 
count_by_time <- function(dataset, date_column, floor_unit) {
    if (!is.character(date_column)) {
        stop("The selected column must be a string.")
    }
  
    if(!(date_column %in% names(dataset))){
        stop("The selected column does not exist in the given dataset.")
    }
  
    if(!(inherits(dataset[[date_column]], 'Date'))){
        stop("The selected column must be a Date.")
    }
  
    if(!(floor_unit %in% c("year", "month", "day"))){
        stop("floor_unit must be one of year, month, or day.")
    }
    
    format_unit <- if(floor_unit == "year") "%Y" else (if(floor_unit == "month") "%b" else "%d") 
    return (dataset %>%
        drop_na(.data[[date_column]]) %>%
        mutate(date_floored = floor_date(.data[[date_column]], unit=floor_unit)) %>%
        mutate(time = format(date_floored, format_unit)) %>%
        group_by(time) %>%
        count(time, name = "count"))
}
```

# Exercise 3
Demonstrate the usage of your function with a few examples. Use one or more new code chunks, describing what you’re doing.

Note: If you want to deliberately show an error, you can use error = TRUE in your code chunk option.

Select Vancouver_trees as the example dataset.
```{r, message = FALSE}
head(vancouver_trees)
```

Now show the total number of trees planted each year
```{r, message = FALSE}
(count_by_time(vancouver_trees, "date_planted", "year"))
```

Show the total number of trees planted each month across all years
```{r, message = FALSE}
(count_by_time(vancouver_trees, "date_planted", "month"))
```

Show the total number of trees planted each day across all months
```{r, message = FALSE}
(count_by_time(vancouver_trees, "date_planted", "day"))
```

The function throws an error when date_column is not a valid string.
```{r, error = TRUE}
date_planted <- as.Date(c("2020-01-01", "2020-01-02"))
(count_by_time(vancouver_trees, date_planted, "year"))
```

It also throws an error when date_column does not exist in the dataset.
```{r, error = TRUE}
(count_by_time(vancouver_trees, "planted_date", "year"))
```

It also throws an error when date_column is not a Date column.
```{r, error = TRUE}
(count_by_time(vancouver_trees, "curb", "year"))
```

It also throws an error when floor_unit is not a value in ("day", "year", "month").
```{r, error = TRUE}
(count_by_time(vancouver_trees, "date_planted", "hour"))
```

# Exercise 4

Running examples is a good way of checking by-eye whether your function is working as expected. But, having a formal “yes or no” check is useful when you move on to other parts of your analysis.

Write formal tests for your function. You should use at least three non-redundant uses of an expect_() function from the testthat package, and they should be contained in a test_that() function (or more than one). They should all pass.

```{r, error = TRUE}
count_by_year <- vancouver_trees %>%
  drop_na(date_planted) %>%
  mutate(date_floored = floor_date(date_planted, unit="year")) %>%
  mutate(time = format(date_floored, "%Y")) %>%
  group_by(time) %>%
  count(time, name = "count")

count_by_month <- vancouver_trees %>%
  drop_na(date_planted) %>%
  mutate(date_floored = floor_date(date_planted, unit="month")) %>%
  mutate(time = format(date_floored, "%b")) %>%
  group_by(time) %>%
  count(time, name = "count")

count_by_day <- vancouver_trees %>%
  drop_na(date_planted) %>%
  mutate(date_floored = floor_date(date_planted, unit="day")) %>%
  mutate(time = format(date_floored, "%d")) %>%
  group_by(time) %>%
  count(time, name = "count")

test_that("Test count_by_time function", {
  # Test if it handles "year" correctly
  expect_equal(count_by_time(vancouver_trees, "date_planted", "year"), count_by_year)
  # Test if it handles "month" correctly
  expect_equal(count_by_time(vancouver_trees, "date_planted", "month"), count_by_month)
  # Test if it handles "day" correctly
  expect_equal(count_by_time(vancouver_trees, "date_planted", "day"), count_by_day)
  # Test if it throws an error when the input column is an object
  expect_error(count_by_time(vancouver_trees, date_planted, "year"), 'The selected column must be a string.')
  # Test if it throws an error when the input column does not exist
  expect_error(count_by_time(vancouver_trees, "planted_date", "year"), 'The selected column does not exist in the given dataset.')
  # Test if it throws an error when the input column is not a Date column
  expect_error(count_by_time(vancouver_trees, "curb", "year"), 'The selected column must be a Date.')
  # Test if it throws an error when the time unit is not valid option
  expect_error(count_by_time(vancouver_trees, "date_planted", "hour"), 'floor_unit must be one of year, month, or day.')
})
```
