---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# count.by.time

<!-- badges: start -->
<!-- badges: end -->

The goal of count.by.time is to provide a function to count the occurrence of data observations by multiple time frames

## Installation

You can install the development version of count.by.time from [GitHub](https://github.com/) with:

```{r}
# install.packages("devtools")
devtools::install_github("stat545ubc-2022/assignment-b1-and-b2-runzw")
```

## Example

This is a basic example that shows you how to count the number of tree planted in different years/months/days based on vancouver_trees dataset.
```{r, message = FALSE}
library(datateachr)
library(dplyr)
library(tidyr)
library(lubridate)
library(count.by.time)
(count_by_time(datateachr::vancouver_trees, "date_planted", "year"))
(count_by_time(datateachr::vancouver_trees, "date_planted", "month"))
(count_by_time(datateachr::vancouver_trees, "date_planted", "day"))
```
